---
title: "Some fun creative title"
author: "Tony Segal, Agustin Fitipaldi, Andrew Reilly, Elijah Stevenson"
date: "June 5, 2025"
format: pdf
---

# Executive Summary

yap here ...

specification below ...

$$
\text{CoreDeposits}_{it} = \beta_0 + \beta_1\text{Crash}_{it} + \gamma_i + \tau_t + \epsilon_{it}
$$
Where $\beta_0$ and $\beta_1$ are the intercept and slope coefficient parameters respectively. $\text{Crash}_{it}$ is a binary variable that is 1 if state i is on the west coast after the Silicon Valley Bank crash. $\gamma_i$ represents state fixed effects and $\tau_t$ stands for time fixed effects. 

Table showing parallel trend assumption below. Lines moving parallel before crash, and then sudden drop in west coast states after crash. ...

```{r}
#| warning: false
#| message: false
#| echo: false

# ========================== GLOBAL CONFIGURATION ==========================
# Edit these parameters to control the entire analysis

# Variable for analysis (both clustering and regression)
analysis_variable <- "Equity Capital to Total Assets - All Institutions"  # Main variable of interest
# Other options: "Equity Capital to Total Assets - All Institutions", 
#                "Net Loans & Leases to Total Assets - All Institutions",
#                "Cost of Funding Earning Assets - All Institutions", etc.

# States to include/exclude
include_california <- FALSE  # Include California in clustering and analysis
exclude_states <- c("Nevada")  # List of states to remove entirely
# Example: c("Nevada", "Delaware", "South Dakota") for multiple states

# Clustering configuration
include_california_in_clustering <- FALSE  # Include CA in the clustering algorithm itself
clustering_variable <- "Equity Capital to Total Assets - All Institutions"  # Variable used for k-means clustering
# Can be different from analysis_variable if desired

print("=== ANALYSIS CONFIGURATION ===")
print(paste("Analysis variable:", analysis_variable))
print(paste("Clustering variable:", clustering_variable))
print(paste("Include California:", include_california))
print(paste("Include CA in clustering:", include_california_in_clustering))
print(paste("Excluded states:", paste(exclude_states, collapse = ", ")))
print("===============================")

# ========================== DATA LOADING AND SETUP ==========================

library(dplyr)
library(readr)
library(gt)
library(ggplot2)
library(estimatr)
library(modelsummary)
library(tidyr)
library(cluster)

lbank <- read_csv("combined_fdic_data_log.csv")

names(lbank)[10] <- "core_deposit"
lbank$State <- as.factor(lbank$State)
lbank$core_deposit <- as.numeric(lbank$core_deposit)
lbank$Date <- as.Date(lbank$Date, format = "%m/%d/%Y")

# Apply global filtering
lbank <- lbank %>%
  filter(!State %in% exclude_states) %>%
  {if(!include_california) filter(., State != "California") else .}

# Ensure data types are correct after filtering
lbank$Date <- as.Date(lbank$Date)  # Ensure Date remains Date type after filtering
lbank$State <- droplevels(lbank$State)  # Remove unused factor levels from excluded states

# Create geographic groupings (conditional on California inclusion)
if(include_california) {
  west_states <- c("California", "Oregon", "Arizona", "Washington", "Idaho", "Utah", "Montana", "Wyoming", "Colorado", "New Mexico")
} else {
  west_states <- c("Oregon", "Arizona", "Washington", "Idaho", "Utah", "Montana", "Wyoming", "Colorado", "New Mexico")
}

lbank$west <- ifelse(lbank$State %in% west_states, 1, 0)
lbank$year <- format(lbank$Date, "%Y")
lbank$post <- ifelse(lbank$Date >= as.Date("2023-03-31"), 1, 0)
lbank$post <- as.factor(lbank$post)
lbank$west <- as.factor(lbank$west)

# Calculate averages for Western and non-Western states using the analysis variable
avg_deposits <- lbank %>%
  group_by(Date, west) %>%
  summarize(avg_analysis_var = mean(get(analysis_variable), na.rm = TRUE)) %>%
  mutate(Region = ifelse(west == 1, "Western States", "Non-Western States"))

# Create single plot with averages
ggplot(data = avg_deposits,
       aes(x = Date, y = avg_analysis_var, color = Region)) +
  geom_line(size = 1.2) +
  geom_vline(xintercept = as.Date("2022-12-31"), color = "red", linetype = "dotted", linewidth = 1) +
  labs(title = paste("Geographic Analysis:", analysis_variable),
       subtitle = "Western vs Non-Western States",
       x = "Date",
       y = paste(analysis_variable, "(log)"),
       color = "Region") + 
  annotate("text", 
           x = as.Date("2022-12-31"), 
           y = min(avg_deposits$avg_analysis_var) + 0.01, 
           label = "Crash", 
           color = "red", 
           vjust = -0.5, 
           angle = 90, 
           size = 4)
```

## Differences-in-Differences Analysis

The graph below shows the difference between western and non-western states' core deposits over time. When the line is above zero, western states are performing better than non-western states. When below zero, non-western states are outperforming western states. The line at zero indicates both regions are changing at the same rate.

```{r}
#| warning: false
#| message: false
#| echo: false

# Calculate difference-in-differences: (change in western) - (change in non-western)

# First create the wide data and UNGROUP it
wide_data <- avg_deposits %>%
  ungroup() %>%  # This is the key - remove any grouping from earlier operations
  select(Date, Region, avg_analysis_var) %>%
  pivot_wider(names_from = Region, values_from = avg_analysis_var) %>%
  arrange(Date)

# Then calculate changes and difference-in-differences using lag
diff_data <- wide_data %>%
  mutate(
    Western_Change = `Western States` - lag(`Western States`),
    NonWestern_Change = `Non-Western States` - lag(`Non-Western States`),
    Difference_in_Differences = Western_Change - NonWestern_Change
  ) %>%
  filter(!is.na(Western_Change))  # Remove first row with NA

# Create differences-in-differences plot
ggplot(data = diff_data, aes(x = Date, y = Difference_in_Differences)) +
  geom_line(size = 1.2, color = "darkblue") +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", alpha = 0.5) +
  geom_vline(xintercept = as.Date("2022-12-31"), color = "red", linetype = "dotted", linewidth = 1) +
  labs(title = paste("Geographic Diff-in-Diff:", analysis_variable),
       subtitle = paste("Change in", analysis_variable, ": (Western Change) - (Non-Western Change)"),
       x = "Date",
       y = "Difference-in-Differences (log)") +
  annotate("text", 
           x = as.Date("2022-12-31"), 
           y = max(diff_data$Difference_in_Differences) - 0.005, 
           label = "Crash", 
           color = "red", 
           vjust = -0.5, 
           angle = 90, 
           size = 4) +
  theme_minimal()
```

## K-Means Clustering Analysis

Instead of using geographic groupings, let's see what happens if we group states based on their banking characteristics using k-means clustering. This data-driven approach might reveal more economically meaningful groupings.

```{r}
#| warning: false
#| message: false
#| echo: false

# Use global configuration parameters
print("=== K-MEANS CLUSTERING SETUP ===")
print(paste("Clustering variable:", clustering_variable))
print(paste("Include CA in clustering:", include_california_in_clustering))

# Prepare clustering data using pre-treatment period (2019-2022)
cluster_data <- lbank %>%
  filter(Date < as.Date("2023-01-01")) %>%  # Pre-treatment period
  {if(!include_california_in_clustering) filter(., State != "California") else .} %>%  # Conditionally remove CA from clustering
  group_by(State) %>%
  summarize(
    clustering_var = mean(get(clustering_variable), na.rm = TRUE)
  ) %>%
  ungroup()

print(paste("Number of states in clustering:", nrow(cluster_data)))

# Convert clustering variable to matrix format for k-means
cluster_vars <- as.matrix(cluster_data$clustering_var)

# Perform k-means clustering with k=2
set.seed(123)
kmeans_result <- kmeans(cluster_vars, centers = 2, nstart = 25)

print(paste("K-means clustering with k=2 using", clustering_variable))
print(table(kmeans_result$cluster))

# Add cluster assignments
cluster_data$cluster <- kmeans_result$cluster
cluster_data$cluster_name <- paste("Cluster", cluster_data$cluster)

# Show cluster assignments
print("K-means cluster assignments:")
cluster_summary <- cluster_data %>%
  arrange(cluster, State) %>%
  select(State, cluster_name, clustering_var)
print(cluster_summary)

# Check which cluster California is in (if included in clustering)
if(include_california_in_clustering) {
  ca_cluster <- cluster_data %>% filter(State == "California") %>% pull(cluster)
  print(paste("California is in Cluster", ca_cluster))
} else {
  print("California was excluded from clustering")
}

# Show cluster characteristics
cluster_stats <- cluster_data %>%
  group_by(cluster_name) %>%
  summarize(
    n_states = n(),
    mean_value = round(mean(clustering_var), 3),
    min_value = round(min(clustering_var), 3),
    max_value = round(max(clustering_var), 3)
  )
print("Cluster characteristics:")
print(cluster_stats)

# Create k-means based diff-in-diff analysis using the analysis variable
# Note: Some states might not have cluster assignments if they weren't in clustering data
lbank_kmeans <- lbank %>%
  left_join(cluster_data %>% select(State, cluster), by = "State") %>%
  filter(!is.na(cluster)) %>%  # Remove states not in clustering
  mutate(cluster_name = paste("Cluster", cluster))

# Calculate averages for each cluster using the analysis variable
avg_deposits_kmeans <- lbank_kmeans %>%
  group_by(Date, cluster_name) %>%
  summarize(avg_analysis_var = mean(get(analysis_variable), na.rm = TRUE)) %>%
  ungroup()

# Create difference-in-differences for k-means groups
wide_data_kmeans <- avg_deposits_kmeans %>%
  pivot_wider(names_from = cluster_name, values_from = avg_analysis_var) %>%
  arrange(Date)

# Get the cluster names dynamically
cluster_cols <- names(wide_data_kmeans)[names(wide_data_kmeans) != "Date"]
cluster1_name <- cluster_cols[1]
cluster2_name <- cluster_cols[2]

diff_data_kmeans <- wide_data_kmeans %>%
  mutate(
    Cluster1_Change = get(cluster1_name) - lag(get(cluster1_name)),
    Cluster2_Change = get(cluster2_name) - lag(get(cluster2_name)),
    Difference_in_Differences = Cluster1_Change - Cluster2_Change
  ) %>%
  filter(!is.na(Cluster1_Change))

# Create comparison plots
# Plot 1: Two-line chart showing levels (like the geographic chart)
ggplot(data = avg_deposits_kmeans, aes(x = Date, y = avg_analysis_var, color = cluster_name)) +
  geom_line(size = 1.2) +
  geom_vline(xintercept = as.Date("2022-12-31"), color = "red", linetype = "dotted", linewidth = 1) +
  labs(title = paste("K-Means Clustering Analysis:", analysis_variable),
       subtitle = paste("Clustered by", clustering_variable),
       x = "Date",
       y = paste(analysis_variable, "(log)"),
       color = "Cluster") + 
  annotate("text", 
           x = as.Date("2022-12-31"), 
           y = min(avg_deposits_kmeans$avg_analysis_var) + 0.01, 
           label = "Crash", 
           color = "red", 
           vjust = -0.5, 
           angle = 90, 
           size = 4) +
  theme_minimal()

# Plot 2: Difference-in-differences
ggplot(data = diff_data_kmeans, aes(x = Date, y = Difference_in_Differences)) +
  geom_line(size = 1.2, color = "darkgreen") +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", alpha = 0.5) +
  geom_vline(xintercept = as.Date("2022-12-31"), color = "red", linetype = "dotted", linewidth = 1) +
  labs(title = paste("K-Means Diff-in-Diff:", analysis_variable),
       subtitle = paste("Clustered by", clustering_variable, ":", cluster1_name, "vs", cluster2_name),
       x = "Date",
       y = "Difference-in-Differences (log)") +
  annotate("text", 
           x = as.Date("2022-12-31"), 
           y = max(diff_data_kmeans$Difference_in_Differences) - 0.005, 
           label = "Crash", 
           color = "red", 
           vjust = -0.5, 
           angle = 90, 
           size = 4) +
  theme_minimal()
```

## Regression Analysis

We'll run two sets of regressions to compare the geographic approach vs the k-means clustering approach.

### Geographic Approach: Western vs Non-Western States

```{r}
#| tbl-cap: "Effect of Silicon Valley Bank Crash on Western States"
#| label: tbl-pannel
#| echo: false

# Use global configuration for regression analysis
print("=== REGRESSION ANALYSIS SETUP ===")
print(paste("Dependent variable:", analysis_variable))
print(paste("States included:", nrow(lbank %>% distinct(State))))

lbank$crash <- ifelse(lbank$west == 1 & lbank$post == 1, 1, 0)

# Create more appropriate time fixed effects for quarterly data
lbank$year <- as.numeric(format(lbank$Date, "%Y"))
lbank$quarter <- quarters(lbank$Date)
lbank$year_quarter <- paste(lbank$year, lbank$quarter, sep = "_")
lbank$year_quarter <- as.factor(lbank$year_quarter)
lbank$State <- as.factor(lbank$State)

print("Treatment variable check:")
print(paste("Crash observations:", sum(lbank$crash)))
print(paste("Post-treatment periods:", length(unique(lbank$Date[lbank$post == 1]))))
print(paste("Western states:", length(unique(lbank$State[lbank$west == 1]))))

# Create formula dynamically based on analysis_variable
# Use year-quarter fixed effects instead of date fixed effects for cleaner quarterly identification
base_formula <- as.formula(paste("`", analysis_variable, "` ~ crash", sep=""))
state_formula <- as.formula(paste("`", analysis_variable, "` ~ crash + State", sep=""))
time_formula <- as.formula(paste("`", analysis_variable, "` ~ crash + year_quarter", sep=""))
full_formula <- as.formula(paste("`", analysis_variable, "` ~ crash + State + year_quarter", sep=""))

#create 4 different models using the analysis variable
model1 <- lm_robust(base_formula, data = lbank, se_type = "HC1")

model2 <- lm_robust(state_formula, data = lbank, se_type = "HC1")

model3 <- lm_robust(time_formula, data = lbank, se_type = "HC1")

model4 <- lm_robust(full_formula, data = lbank, se_type = "HC1")

#create list of models
models_list <- list(
  "(1)" = model1,
  "(2)" = model2,
  "(3)" = model3,
  "(4)" = model4
)

modelsummary(models_list,
             title = paste("Effect of Silicon Valley Bank Crash on", analysis_variable),
             stars = TRUE,
             gof_map = c("nobs"),
             coef_map = c("crash" = "Bank Crash",
                          "(Intercept)" = "Intercept"),
             coef_omit = "State|year_quarter",
             fmt = "%.3f", 
             notes = "Robust standard errors in parenthesis.",
             add_rows = tribble(
               ~term,         ~"(1)",  ~"(2)",  ~"(3)",  ~"(4)",
               "State FE",    "No",    "Yes",   "No",    "Yes",
               "Quarter FE",  "No",    "No",    "Yes",   "Yes"
             ),
             output = "gt") %>% 
  tab_spanner(
    label = analysis_variable,
    columns = everything()
  ) %>% 
  tab_options(
    table.width = pct(90),
    heading.align = "left"
  )
```

### K-Means Clustering Approach

Now let's run the same regression models but using the k-means cluster assignments instead of geographic groupings.

```{r}
#| tbl-cap: "Effect of Silicon Valley Bank Crash using K-Means Clusters"
#| label: tbl-kmeans
#| echo: false

# Create k-means treatment variable
# Only include states that were part of the clustering
lbank_reg_kmeans <- lbank %>%
  left_join(cluster_data %>% select(State, cluster), by = "State") %>%
  filter(!is.na(cluster)) %>%  # Only states included in clustering
  mutate(
    kmeans_treatment = ifelse(cluster == 1, 1, 0),  # Treat cluster 1 as "treatment"
    kmeans_crash = ifelse(kmeans_treatment == 1 & post == 1, 1, 0)
  )

# Create time fixed effects for k-means regression too
lbank_reg_kmeans$year <- as.numeric(format(lbank_reg_kmeans$Date, "%Y"))
lbank_reg_kmeans$quarter <- quarters(lbank_reg_kmeans$Date)
lbank_reg_kmeans$year_quarter <- paste(lbank_reg_kmeans$year, lbank_reg_kmeans$quarter, sep = "_")
lbank_reg_kmeans$year_quarter <- as.factor(lbank_reg_kmeans$year_quarter)
lbank_reg_kmeans$State <- as.factor(lbank_reg_kmeans$State)

print("=== K-MEANS REGRESSION SETUP ===")
print(paste("Cluster 1 states:", sum(lbank_reg_kmeans$kmeans_treatment == 1 & !duplicated(lbank_reg_kmeans$State))))
print(paste("Cluster 2 states:", sum(lbank_reg_kmeans$kmeans_treatment == 0 & !duplicated(lbank_reg_kmeans$State))))

# Create formulas for k-means approach
# Wrap variable name in backticks to handle spaces and special characters
base_formula_kmeans <- as.formula(paste("`", analysis_variable, "` ~ kmeans_crash", sep=""))
state_formula_kmeans <- as.formula(paste("`", analysis_variable, "` ~ kmeans_crash + State", sep=""))
time_formula_kmeans <- as.formula(paste("`", analysis_variable, "` ~ kmeans_crash + year_quarter", sep=""))
full_formula_kmeans <- as.formula(paste("`", analysis_variable, "` ~ kmeans_crash + State + year_quarter", sep=""))

# Create k-means models
model1_kmeans <- lm_robust(base_formula_kmeans, data = lbank_reg_kmeans, se_type = "HC1")
model2_kmeans <- lm_robust(state_formula_kmeans, data = lbank_reg_kmeans, se_type = "HC1")
model3_kmeans <- lm_robust(time_formula_kmeans, data = lbank_reg_kmeans, se_type = "HC1")
model4_kmeans <- lm_robust(full_formula_kmeans, data = lbank_reg_kmeans, se_type = "HC1")

# Create list of k-means models
models_list_kmeans <- list(
  "(1)" = model1_kmeans,
  "(2)" = model2_kmeans,
  "(3)" = model3_kmeans,
  "(4)" = model4_kmeans
)

modelsummary(models_list_kmeans,
             title = paste("Effect of Silicon Valley Bank Crash on", analysis_variable, "- K-Means Clusters"),
             stars = TRUE,
             gof_map = c("nobs"),
             coef_map = c("kmeans_crash" = "K-Means Cluster Crash",
                          "(Intercept)" = "Intercept"),
             coef_omit = "State|year_quarter",
             fmt = "%.3f", 
             notes = "Robust standard errors in parenthesis. Cluster 1 treated as treatment group.",
             add_rows = tribble(
               ~term,         ~"(1)",  ~"(2)",  ~"(3)",  ~"(4)",
               "State FE",    "No",    "Yes",   "No",    "Yes",
               "Quarter FE",  "No",    "No",    "Yes",   "Yes"
             ),
             output = "gt") %>% 
  tab_spanner(
    label = paste(analysis_variable, "- K-Means Approach"),
    columns = everything()
  ) %>% 
  tab_options(
    table.width = pct(90),
    heading.align = "left"
  )
```

Can introduce these tables and graphs before if necessary:

```{r}
#| warning: false
#| message: false
#| echo: false
#| tbl-cap: "Log Core Deposits to Total Liabilities"
#| label: tbl-log_core

# Use the already filtered lbank data from global configuration
# Ensure Date is properly formatted (may have been converted to factor during filtering)
bank <- lbank %>% 
  mutate(
    Date = as.Date(Date),  # Ensure Date is Date type, not factor
    year = as.numeric(format(Date, "%Y")),
    month = as.numeric(format(Date, "%m"))) 

#get 2022 Q4 data for analysis variable by State
q4_2022 <- bank %>% 
  filter(year == 2022, month %in% c(10, 11, 12)) %>% 
  group_by(State) %>% 
  summarize(total_analysis_var = sum(get(analysis_variable), na.rm = TRUE))

#get 2023 Q1 data for analysis variable by State
q1_2023 <- bank %>% 
  filter(year == 2023, month %in% c(1, 2, 3)) %>% 
  group_by(State) %>% 
  summarize(total_analysis_var = sum(get(analysis_variable), na.rm = TRUE))

#merge dataframes together for table
merged <- merge(q4_2022, q1_2023, by = "State", suffixes = c("_q4_2022", "_q1_2023"))

#calculate difference between 2022 and 2023
merged$diff <- merged$total_analysis_var_q1_2023 - merged$total_analysis_var_q4_2022   

#sort descending order
merged <- merged %>% 
  arrange(diff)

#construct gt table
gt_table <- head(merged, 10) %>% 
  gt() %>% 
  cols_label(
    total_analysis_var_q4_2022 = "2022 Q4",
    total_analysis_var_q1_2023 = "2023 Q1",
    diff = paste("Change in", analysis_variable, "After Shock")
  ) %>% 
  cols_align(
    align = "left",
    columns = State
  ) %>% 
  cols_align(
    align = "center",
    columns = c(total_analysis_var_q4_2022, total_analysis_var_q1_2023, diff)
  )
  
# print table 
gt_table

# Add bar graph
ggplot(head(merged, 10), aes(x = reorder(State, diff), y = diff)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = paste("Change in", analysis_variable, "After Shock"),
       x = "State",
       y = paste("Difference in", analysis_variable, "(2022 Q4 - 2023 Q1)")) +
  theme_minimal()
```

Ill try to add comments to the code later too